Replacing standard galaxy profiles with \\ mixtures of Gaussians

### 2013-03-10 documents/mixtures/mixture_models.tex
cites and acks

Not only does a sampling representation of the profiles require many
more compute operations per render than a MoG representation, it also
requires more code complexity.  For the sampling to be both accurate
and efficient, it has to be strongly adaptive (with much higher
density in the center).  This adds complexity and choices into the
code, leading to engineering and software maintenance and testing
overheads that cost not just compute time but human time.  One
response to this is to build large numbers of pre-computed profile PSF
convolutions and interpolate among them.  This is the strategy taken
in the \project{SDSS} pipelines (\citealt{lupton}), but that requires
even \emph{more} human overhead, because not only does the sampling
and convolution have to be done correctly, so does the numerical
interpolation.  Another less common strategy is to find other kinds of
function expansions (\citealt{bendi}).  Finally, a sampling
approximation---or any finite numerical approximation---is effectively
something like a ``mixture of delta functions'' or ``mixture of
integration kernels'' approximation.  Any such approximation (at fixed
accuracy) will have far more degrees of freedom (complexity) than the
MoG approximations given below.

@@

\acknowledgements It is a pleasure to thank
      Brendon Brewer (Auckland),
      Jim Bosch (Princeton),
      Kevin Bundy (IPMU),
      and an anonymous referee
for valuable comments.  This work was supported in part by NASA (grant
NNX12AI50G) and the NSF (grant IIS-1124794).  All the results shown in
the figures and tables and more are available in machine-readable form
from author DWH upon request.

\begin{thebibliography}{70}
\bibitem[Bendinelli(1991)]{bendinelli}
Bendinelli,~O., 1991, \apj, 366, 599
\bibitem[Bendinelli et al.(1993)]{bendi}
Bendinelli,~O., Ciotti,~L., \& Parmeggiani,~G.\ 1993, \aap, 279, 668 
\bibitem[Bendinelli \& Parmeggiani(1995)]{bendinelli2}
Bendinelli,~O. \& Parmeggiani,~G., 1995, \aj, 109, 572
\bibitem[Bosch(2010)]{bosch}
Bosch,~J., 2010, \aj, 140, 870
\bibitem[Bovy \etal(2011a)]{xd}
Bovy,~J., Hogg,~D.~W., \& Roweis,~S., 2011, Ann. Appl. Stat., 5, 1657
\bibitem[Bovy \etal(2011b)]{xdqso}
Bovy,~J., \etal, 2011, \apj, 729, 141
\bibitem[Bovy \etal(2012)]{xdqsoz}
Bovy,~J. \etal, 2012, \aj, 749, 41
\bibitem[Bundy \etal(2012)]{bundy}
Bundy,~K. \etal, \aj, in press
\bibitem[Cappellari(2002)]{cappellari}
Cappellari,~M., 2002, \mnras, 333, 400
\bibitem[Ciotti \& Bertin(1999)]{ciotti}
Ciotti,~L. \& Bertin,~G., 1999, \aa, 352, 447
\bibitem[de~Vaucouleurs(1948)]{dev}
de~Vaucouleurs,~G., 1948, Annales d'Astrophysique, 11, 247
\bibitem[Emsellem \etal(1994)]{emsellem}
Emsellem,~E., Monnet,~G., Bacon,~R., \& Nieto,~J.-L., 1994, \aap, 285, 739
\bibitem[Kochanek et al.(2000)]{kochanek}
Kochanek,~C.~S., Falco,~E.~E., Impey,~C.~D., et al., 2000, \apj, 535, 692
\bibitem[Lupton \etal(2001)]{lupton}
Lupton,~R., Gunn,~J.~E., Ivezic,~Z., Knapp,~G.~R., Kent,~S.~M., \& Yasuda,~N., 2001, ASPC, 238, 269
\bibitem[Peng \etal(2002)]{galfit}
Peng,~C.~Y., Ho, L.~C., Impey,~C.~D., \& Rix,~H.-W., 2002, \aj, 124, 266
\bibitem[S{\'e}rsic(1963)]{ser}
S{\'e}rsic,~J.~L., 1963, Boletin de la Asociacion Argentina de Astronomia La Plata Argentina, 6, 41
\bibitem[Shaw(2012)]{shaw}
Shaw,~R.~A., ed., 2012, LSST Data Challenge Handbook (Version 2.0; Tucson, AZ: LSST Corp.)
\bibitem[Simard et al.(2011)]{bulgedisk}
Simard,~L., Mendel,~J.~T., Patton,~D.~R., Ellison,~S.~L., \& McConnachie,~A.~W., 2011, \apjs, 196, 11
\bibitem[Trujillo \etal(2001)]{trujillo}
Trujillo,~I., Aguerri,~J.~A.~L., Cepa,~J., \& Guti{\'e}rrez,~C.~M., 2001, \mnras, 321, 269 
\bibitem[van~den~Bosch et al.(2008)]{vdb}
van~den~Bosch,~R.~C.~E., van de Ven,~G., Verolme,~E.~K., Cappellari,~M., \& de~Zeeuw,~P.~T., 2008, \mnras, 385, 647
\end{thebibliography}

>>>>

Not only does a sampling representation of the profiles require many
more compute operations per render than a MoG representation, it also
requires more code complexity.  For the sampling to be both accurate
and efficient, it has to be strongly adaptive (with much higher
density in the center).  This adds complexity and choices into the
code, leading to engineering and software maintenance and testing
overheads that cost not just compute time but human time.  One
response to this is to build large numbers of pre-computed profile PSF
convolutions and interpolate among them.  This is the strategy taken
in the \project{SDSS} pipelines (\citealt{lupton}), but that requires
even \emph{more} human overhead, because not only does the sampling
and convolution have to be done correctly, so does the numerical
interpolation.  Another less common strategy is to find other kinds of
function expansions or approximations (\citealt{bendi},
\citealt{spergel}).  Finally, a sampling approximation---or any finite
numerical approximation---is effectively something like a ``mixture of
delta functions'' or ``mixture of integration kernels'' approximation.
Any such approximation (at fixed accuracy) will have far more degrees
of freedom (complexity) than the MoG approximations given below.

@@

\acknowledgements It is a pleasure to thank
      Brendon Brewer (Auckland),
      Jim Bosch (Princeton),
      Kevin Bundy (IPMU),
      Luca Ciotti (Bologna),
      Rachel Mandelbaum (CMU),
      Chien Peng (OCIW),
      and an anonymous referee
for valuable comments.  This work was supported in part by NASA (grant
NNX12AI50G) and the NSF (grant IIS-1124794).  All the results shown in
the figures and tables and more are available in machine-readable form
from author DWH upon request.

\begin{thebibliography}{70}
\bibitem[Bendinelli(1991)]{bendinelli}
Bendinelli,~O., 1991, \apj, 366, 599
\bibitem[Bendinelli et al.(1993)]{bendi}
Bendinelli,~O., Ciotti,~L., \& Parmeggiani,~G.\ 1993, \aap, 279, 668 
\bibitem[Bendinelli \& Parmeggiani(1995)]{bendinelli2}
Bendinelli,~O. \& Parmeggiani,~G., 1995, \aj, 109, 572
\bibitem[Bosch(2010)]{bosch}
Bosch,~J., 2010, \aj, 140, 870
\bibitem[Bovy \etal(2011a)]{xd}
Bovy,~J., Hogg,~D.~W., \& Roweis,~S., 2011, Ann. Appl. Stat., 5, 1657
\bibitem[Bovy \etal(2011b)]{xdqso}
Bovy,~J., \etal, 2011, \apj, 729, 141
\bibitem[Bovy \etal(2012)]{xdqsoz}
Bovy,~J. \etal, 2012, \aj, 749, 41
\bibitem[Bundy \etal(2012)]{bundy}
Bundy,~K. \etal, \aj, in press
\bibitem[Cappellari(2002)]{cappellari}
Cappellari,~M., 2002, \mnras, 333, 400
\bibitem[Ciotti \& Bertin(1999)]{ciotti}
Ciotti,~L. \& Bertin,~G., 1999, \aa, 352, 447
\bibitem[de~Vaucouleurs(1948)]{dev}
de~Vaucouleurs,~G., 1948, Annales d'Astrophysique, 11, 247
\bibitem[Emsellem \etal(1994)]{emsellem}
Emsellem,~E., Monnet,~G., Bacon,~R., \& Nieto,~J.-L., 1994, \aap, 285, 739
\bibitem[Kochanek et al.(2000)]{kochanek}
Kochanek,~C.~S., Falco,~E.~E., Impey,~C.~D., et al., 2000, \apj, 535, 692
\bibitem[Lupton \etal(2001)]{lupton}
Lupton,~R., Gunn,~J.~E., Ivezic,~Z., Knapp,~G.~R., Kent,~S.~M., \& Yasuda,~N., 2001, ASPC, 238, 269
\bibitem[Peng \etal(2002)]{galfit}
Peng,~C.~Y., Ho, L.~C., Impey,~C.~D., \& Rix,~H.-W., 2002, \aj, 124, 266
\bibitem[S{\'e}rsic(1963)]{ser}
S{\'e}rsic,~J.~L., 1963, Boletin de la Asociacion Argentina de Astronomia La Plata Argentina, 6, 41
\bibitem[Shaw(2012)]{shaw}
Shaw,~R.~A., ed., 2012, LSST Data Challenge Handbook (Version 2.0; Tucson, AZ: LSST Corp.)
\bibitem[Simard et al.(2011)]{bulgedisk}
Simard,~L., Mendel,~J.~T., Patton,~D.~R., Ellison,~S.~L., \& McConnachie,~A.~W., 2011, \apjs, 196, 11
\bibitem[Spergel(2010)]{spergel}
Spergel,~D.~N.\ 2010, \apjs, 191, 58
\bibitem[Trujillo \etal(2001)]{trujillo}
Trujillo,~I., Aguerri,~J.~A.~L., Cepa,~J., \& Guti{\'e}rrez,~C.~M., 2001, \mnras, 321, 269 
\bibitem[van~den~Bosch et al.(2008)]{vdb}
van~den~Bosch,~R.~C.~E., van de Ven,~G., Verolme,~E.~K., Cappellari,~M., \& de~Zeeuw,~P.~T., 2008, \mnras, 385, 647
\end{thebibliography}
